---
title: Juvenile Snowshoe Hares in Bonanza Creek Experimental Forest
author: "Devon Rossi"
date: "11/7/2021"
output: html_document
---
```{r setup, include= FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(lubridate)
library(kableExtra)
library(dplyr)
library(broom)
library(ggplot2)
library(effsize)
library(equatiomatic)
library(ggbeeswarm)
library(stargazer)
#Attach packages
```

### 1. Overview 
This report provides an exploratory overview of snowshoe hares in Bonanza Creek Experimental Forest. Within this report, data visualization for annual juvenile hare trap counts calculated from 1999 - 2012 and for the difference between juvenile hare weights by sex and site (Bonanza Riparian, Bonanza Mature, Bonanza Black Spruce) are included. A summary statistics table is provided of juvenile hare weight comparison to understand the differences in mean, median, and standard deviation of weights between male and female juvenile snowshoe hares.

#### Data citation
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14


### 2. Data and Methods 
Bonanza snowshoe hare data related to traps, weight, sex, and site are collected from Kielland, Chapin, and Ruess of the Bonanza Creek LTER. Following data visualization and exploratory histograms and QQ-plots, a two-sample t-test was calculated using a significance level ($\alpha$) of 0.05 throughout. Differences between male and female snowshoe hares are described by Cohen's *d* effect size. The relationship between juvenile hare snowshoe hindfoot length and weight is explored by simple linear regression.

### 3. Juvenile Hare Initial Investigation
```{r}
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv")) %>% 
  mutate(sex = case_when(
    sex == "m" ~ "Male",
    sex == "f" ~ "Female"))
```


## A. Data Visualization: Annual juvenile hare trap counts 
```{r}
juvenile_hares <- bonanza_hares %>% 
  filter(age == "j") %>% #filter for juvenile hares
  mutate(date_new = mdy(date)) %>% #new column for date
  mutate(year = year(date_new)) %>% #new column for year
  group_by(year)

j_traps <- juvenile_hares %>% 
  summarize(count = n())
```

```{r}
juvenile_hares_count <- juvenile_hares %>% 
  select(year, sex, weight) %>% 
  group_by(sex) %>% 
  summarize(mean = mean(weight, na.rm = TRUE),
            median = median(weight, na.rm = TRUE),
            sd = sd(weight, na.rm = TRUE),
            max = max(weight, na.rm = TRUE),
            min = min(weight, na.rm = TRUE),
            count = n(), na.rm = TRUE)
#find mean, median, and sd for juvenile hare count

```

```{r}
ggplot(data = j_traps, aes(x = year, y = count)) +
  geom_line(color = "orange", fill = "orange4", size = 1.5) +
  theme_light() +
  scale_x_continuous(breaks = 1999:2012) +
  labs(title = "Annual Juvenile Hare Trap Counts",
       x = "Year",
       y = "Trap Counts",
       caption = "Figure 1: Juvenile hare trap counts from 1999 - 2012") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(face = "bold", size = 12)) +
  theme(axis.title = element_text(face = "bold", size = 10)) +
  theme(plot.background = element_rect(fill = "white"))
  
```


There were 2 juvenile hare traps in 2010 and 126 traps in 1999, illustrating the minimum and maximum juvenile hare traps within Bonanza Creek Experimental Forest respectively.
The mean trap count is `r mean(j_traps$count)` and the median trap count is `r median(j_traps$count)`.There is a significant decrease between trap count in 1999 to 2012. As the count is impacted by the number of days and the number of traps in the study each year, we must consider standardizing the juvenile hare population through refining the number and days in which traps are set up. For example, setting up the same number of traps each year on the same days will result in greater overall standardization.


## B. Data Visualization: Juvenile Hare Weights by Sex and Site
```{r}
juvenile_hare_weights <- juvenile_hares %>% 
  select(weight, sex, grid) %>% 
  mutate(grid = case_when(
    grid == "bonbs" ~ "Bonanza Black Spruce",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonrip" ~ "Bonanza Riparian"))

ggplot(data = juvenile_hare_weights, aes(x = sex, y = weight)) +
  geom_jitter(aes(color = sex), size = 1.5) +
  geom_boxplot(fill = NA) +
  scale_color_manual(values = c("darkorchid4", "darkorange1")) +
  theme_light() +
  theme(legend.position = "none") +
  facet_wrap(~ grid) + #remember to connect ggplot to labs with '+'
  labs(x = "Sex",
       y = "Weight (kg)",
       caption = "Figure 2: Juvenile hare weight comparison by sex and site") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(family = "Times New Roman", face = "bold", size = 12)) +
  theme(axis.title = element_text(face = "bold", size = 10))

```


There is little difference between mean and median weights throughout each site. The difference between median weight of male and female juvenile snowshoe hares is the greatest at the Riparian and Mature sites, respectively. 

There is little difference in mean and median weights by sex at the Bonanza Black Spruce site. It is important to note, however, that there are significantly less samples from this site. Increased sample count of male and female hares is required to determine the actual difference in means at this site.

## C. Juvenile Snowshoe Hares: Weight comparison by sex
```{r}
juvenile_hare_weight_comparison <- juvenile_hares %>% 
  select(sex, weight) %>% #select sex & weight
  group_by(sex) %>% 
  drop_na(weight) %>% 
  summarize(mean_weight = round(mean(weight), 2),
        median_weight = round(median(weight), 2),
        sd_weight = round(sd(weight), 2),
        n = n()) #how many are there

#create summary table for weight comparison
juvenile_hare_weight_comparison %>% 
  kable(col.names =c("Sex",
                     "Mean",
                     "Median",
                     "Standard Deviation",
                     "Count"),
caption = "<center> Table 1: Comparative statistics comparing weight between male and female juvenile snowshoe hares") %>% 
  kable_styling(full_width = FALSE, font = 12)
```

```{r include = FALSE}
#Exploratory histogram to determine normality
ggplot(data = juvenile_hares, aes(x = weight)) +
  geom_histogram(bins = 12) +
  facet_wrap(~sex)
```

```{r include = FALSE}
#Exploratory qqplot to determine normality
ggplot(data = juvenile_hares, aes(sample = weight)) +
  geom_qq() +
  facet_wrap(~sex)
```

```{r}
#T-Test

female_hare_weight <- juvenile_hares %>% #female weight vector
  filter(sex == "Female") %>% 
  drop_na(weight) %>% #drop na weights
  pull(weight)
  

male_hare_weight <- juvenile_hares %>% #male weight vector
  filter(sex == "Male") %>% 
  drop_na(weight) %>% #drop na weights
  pull(weight)

# Get the means & SDs for each: 
male_hare_mean <- mean(male_hare_weight)
female_hare_mean <- mean(female_hare_weight)

male_hare_sd <- sd(male_hare_weight)
female_hare_sd <- sd(female_hare_weight)

# Get the effect size: 
hare_effsize <- cohen.d(male_hare_weight, female_hare_weight)

#Run the t-test
hare_t <- t.test(male_hare_weight, female_hare_weight) 

# Get the tidy model results to call outputs in-line: 
hare_ttest_tidy <- tidy(hare_t)
```


```{r}
# Cohen's d effect size
hare_cohen <- cohen.d(female_hare_weight, male_hare_weight)
```
On average, male hare vs. female hare weight (`r round(male_hare_mean,2)` $\pm$ `r round(male_hare_sd, 2)` and `r round(female_hare_mean,2)` $\pm$ `r round(female_hare_sd, 2)` mm, respectively; mean $\pm$ 1 standard deviation).

While the absolute difference in means is `r round(male_hare_mean, 2) - round(female_hare_mean, 2)` mm (a `r round((male_hare_mean - female_hare_mean)/((male_hare_mean + female_hare_mean)/2)*100, 2)`% difference), the difference in means is significant (Welch's two-sample t-test: t(`r round(hare_ttest_tidy$parameter,2)`) = `r round(hare_ttest_tidy$statistic,2)`, p < 0.001), and the effect size is small (Cohen's *d* = `r round(hare_effsize$estimate,2)`).

## D. Linear Regression: Exploring Relationship between juvenile weight & hind foot length 
```{r}
ggplot(data = juvenile_hares, aes(x = hindft, y = weight)) +
  geom_point(aes(color= sex, pch = sex)) +
  geom_smooth(method = "lm") +
  theme_light() +
  labs(title = "Hind Foot Length and Weight",
       x = "Hind Foot (mm)",
       y = "Weight (kg)",
       caption = "Figure 3: Hind Foot Length and Weight Comparison by Sex") +
  theme(plot.title = element_text(hjust = 0.5)) +
  theme(plot.title = element_text(family = "Times New Roman", face = "bold", size = 12)) +
  theme(axis.title = element_text(face = "bold", size = 10))

```


```{r include = FALSE}
# Simple linear regression

hare_lm <- lm(hindft ~ weight, data = juvenile_hares)


hare_lm_tidy <- tidy(hare_lm)
hare_lm_glance <- glance(hare_lm)


extract_eq(model = hare_lm, use_coefs = TRUE)
plot(hare_lm)

hare_cor <- cor.test(juvenile_hares$hindft, juvenile_hares$weight)

# Tidy version of correlation output:
hare_cor_tidy <- tidy(hare_cor)

```


The relationship between snowshoe hare hind foot length and weight are linear between both male and female populations (Figure 3). Simple linear regression illustrates that hind foot length significantly predicts snowshoe hare weight (kg) (p < 0.001, R^2^ = `r round(hare_lm_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(hare_lm_tidy$estimate[2], 2)` g mm^-1^ (i.e., for each one millimeter increase in hind foot length we expect an average increase in weight `r round(hare_lm_tidy$estimate[2], 2)` g). 

Hind foot length and weight have a strong, positive correlation (Pearson's *r* = `r round(hare_cor_tidy$estimate,2)`, p < 0.001). Exploratory histogram and QQ plots (not included) reveal normally distributed and homoscedastic residuals. 

## Summary
- Juvenile hare trap count decreases from 1999 - 2012, with a max of 126 in 1999 and a min of 2 in 2012
- Juvenile hare weights by sex and site illustrates little difference in median weight depending on site
- Male and female snowshoe hares show 10% significance in difference between mean weights 
- Linear regression model illustrates strong positive correlation between hind foot length and hare weight (exploratory histogram and QQ plots confirm normal distribution)

#### Data citation: 
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

