---
title: Juvenile snowshoe hares in Bonanza Creek Experimental Forest
author: "Devon Rossi"
date: "11/7/2021"
output: html_document
---

```{r setup, include= TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
```

### 1. Overview
Introduction: A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

### 2. Data and Methods
Data and analyses: A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

### 3. Preliminary results: Juvenile Hare Initial Investigation

#### Read in bonanza hares data
```{r}
bonanza_hares <- read_csv(here("data", "bonanza_hares.csv"))
```

#### A. Annual juvenile hare trap counts
Count the total number of juvenile hare trappings during each year of the study, and create a finalized data visualization of the counts by year. Include a figure caption below the figure in your report. 
```{r}



```

What were the minimum and maximum counts of juvenile hare trappings?
Mean or median annual number of juvenile hares trapped?
General/notable trends?
These are absolute counts, but this count is also impacted by effort (e.g. the number of days and number of traps in the study each year). What would you suggest be done moving forward (you are not expected to actually do this) to standardize the juvenile hare population in future analyses?

#### B. Visualize juvenile hare weights
```{r}







```
Then, in 2 - 3 sentences below your visualization(s) describe in text the main takeaways that you would want the audience to understand from your figures.

#### C. Juvenile weight comparison (male & female snowshoe hares)
```{r}







```
A 2 - 3 sentence comprehensive statement of the means comparison, including at least:
The actual difference in means
Effect size (Cohen's *d*) (see effsize::cohen.d() example in the Lab 5 key)
Outcome of an unpaired two-sample t-test

#### D. Relationship between juvenile weight & hind foot length
```{r}







```
In 2 - 3 sentences below your figure, describe the outcome of your linear regression exploration (e.g. slope interpretation, R2 value, and Pearson's r correlation) and touch briefly on which assumption(s) of linear regression may be a concern, and additional thoughts you have about how well (or not well…) the linear model describes the relationship. What worries you? 

## Summary
Include a very short summary describing your main findings from your juvenile hares exploration (bullet points are fine)


## Citations
Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

